---
title: 'Data Visualization Practice'
author: ""
output: html_document
---

```{r message = FALSE, echo=FALSE}
library(tidyverse)
library(janitor)
bike <- read_csv("https://raw.githubusercontent.com/COSMOS-DataScience/slides/main/week01/data/metro-bike-share-trips.csv")
```

## Question 1

Look at the distribution of duration of bike rides with a histogram. 
Is the distribution symmetric or skewed? If skewed, to which direction is it skewed?
In such distributions would you expect mean > median or median < mean? 
Calculate the mean and median to confirm.


```{r echo=FALSE, message=FALSE}
#generate graph code here


#calculate mean duration and median duration here
mean_dur = NA

med_dur = NA

```


We create a new variable called `standard_bike`. This variable indicates whether a bike is standard or not. Make sure to "save" this variable to your data  frame.

```{r echo=FALSE, message=FALSE}
bike <- 
  bike %>%
    mutate(standard_bike = case_when(
             bike_type == "standard" ~ TRUE, 
             bike_type != "standard" ~ FALSE))

glimpse(bike)
```


## Question 2

Make a visual comparing the duration of bike rides for standard and non-standard bikes.
```{r echo=FALSE, message=FALSE}
#Use an appropriate graph to compare distribution of bike ride durations

```


It seems like there are some bike rentals that are 1440 minutes long which is basically a day. How many of the bike rentals that took the whole day? Answer with inline code.

```{r echo=FALSE, message=FALSE}
whole_day_dur<-bike %>% 
  filter(duration >= 1440) %>% 
  nrow()
```

`r whole_day_dur` bike rentals took the whole day.

We create a new variable called `day_long` this variable should show whether a bike rental is day long (1440 minutes) or not. Make sure to save this variable in your data frame.

```{r echo=FALSE, message=FALSE}
bike <- 
  bike %>%
    mutate(day_long = case_when(
             duration >= 1440 ~ TRUE, 
             duration < 1440 ~ FALSE))

glimpse(bike)
```

Change the type of the `passholder_type` to a correct type and save it in the data frame accordingly. 

```{r}
bike <- bike %>% 
  mutate(passholder_type = as.factor(passholder_type))
```

## Question 3

Make a visual the relationship between `passholder_type` and `day_long`

```{r echo = FALSE, message = FALSE, warning = FALSE}

```
