---
title: "Working with R"
author: "Brian Schetzsle"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css", "https://cdnjs.cloudflare.com/ajax/libs/font-awesome/5.15.3/css/all.min.css"]
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "cols_macro.js"
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"

---

class: title-slide

```{r echo = FALSE, warning=FALSE}
library(fabricerin)
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

Adapted from slides by Mine Dogucu
]

---

## Reminder

- Close all apps on your computer.  
- Open slides for this session from the cluster website (https://cosmos-datascience.netlify.app). 

---

# Introductions

---

# First things first:
- Do you have a computer?
- Do you have R on your computer?
- Do you have RStudio on your computer?

---

# A Walkthrough of RStudio
## Why it's a nice environment for R coding

---

# A Walkthrough of RStudio
## Create a new document in which to do R coding

---

class: inverse middle center

.font100[hello woRld]

---

class: center

<video width="80%" height="50%" align = "center" controls>
  <source src="screencast/01a-hello-world.mp4" type="video/mp4">
</video>

.footnote[Created by Dr. Mine Dogucu]

---

## Practice

- open RStudio.
- write the R code you have just seen.
- help each other with problems.
- Explore the relationships between the different quadrants in RStudio
- Create an R document, type code there and execute it in the console using Command+Enter

---

# R review
## Object assignment operator

```{r}
birth_year <- 1950
```

---

# R review

R is case-sensitive


```{r error=TRUE}

my_age <- 2022 - birth_year

My_age
```



---

## Variables in R

`character`: takes string values (e.g. a person's name, address)

--

`integer`: integer (single precision)  

--

`double`: floating decimal (double precision) 

--

`numeric`: integer or double  

--

`factor`: categorical variables with different levels  

--

`logical`: TRUE (1), FALSE (0)  



---


# R review

If something comes in quotes, it is not defined in R. 


```{r}

ages <- c(25, my_age, 32)

names <- c("Menglin", "Mine", "Rafael")

data.frame(age = ages, name = names)
```

---

# R review
## Functions and Arguments

```{r eval=FALSE}
do(something)
```

`do()` is a function;   
`something` is the argument of the function.

--

```{r eval=FALSE}
do(something, colorful)
```

`do()` is a function;   
`something` is the first argument of the function;   
`colorful` is the second argument of the function.

---

# R review

## Getting Help

In order to get any help we can use `?` followed by function (or object) name. 

```{r eval=FALSE}
?c
```


---

# Some basic data structures
- variables/assignment
- combine/vector
- matrix

```{r eval=FALSE}
temp_variable = 1
temp_vector = c(1, 2, 3)
temp_matrix = matrix(c(1,2,3,4), nrow=2, ncol=2, 
                     byrow=TRUE)
```
