---
title: "Introduction to tidyverse"
author: "Zahra Moslemi"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css", "https://cdnjs.cloudflare.com/ajax/libs/font-awesome/5.15.3/css/all.min.css"]
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "cols_macro.js"
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"

---

class: title-slide

```{r echo = FALSE, warning=FALSE}
library(fabricerin)

```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

Adapted from slides by Mine Dogucu
]



---

## The tidyverse style guide

>canyoureadthissentence?  

.important[
Good coding style is like correct punctuation: you can manage without it, but it makes things easier to read.
]


- The most important thing about tidyverse style guide is that it provides consistency, making code easier to write because you need to make fewer decisions.


---
- Avoid using special characters in file names - stick with numbers, letters, -, and _.

```{r, message=FALSE, eval = FALSE}
# Good
fit_models.R
utility_functions.R

# Bad
fit models.R
foo.r
stuff.r
```
---

- Put a space after a comma, not before.

```{r, message=FALSE, eval = FALSE}
# Good
age <- c(6, 9, 15)
# Bad
age <- c(6,9,15)
```

- Object names are all lower case, with words separated by an underscore.
- Before and after operators (e.g. <-, =) leave spaces.

```{r, message=FALSE, eval = FALSE}
# Good
data.frame(age_kid = age)

# Bad
data.frame(Age_kid=age)
```

-[The tidyverse style guide](https://style.tidyverse.org/)
---
class: inverse middle
.font50[The Pipe Operator]

```{r echo = FALSE}
library(magrittr)
```


---

class: middle
>Three solutions to a single problem

What is the average of 4, 8, 16 approximately?

---

class: middle

1.What is the average of **4, 8, 16** approximately?

---

class: middle

2.What is the **average** of 4, 8, 16 approximately?

---

class: middle

3.What is the average of 4, 8, 16 **approximately**?

---

class: middle inverse

.font50[Solution 1: Functions within Functions]

---

```{r}
c(4, 8, 16)
```

--

<hr>

```{r}
mean(c(4, 8, 16))
```

--

<hr>

```{r}
round(mean(c(4, 8, 16)))
```

---

class: middle

**Problem with writing functions within functions**

Things will get messy and more difficult to read and debug as we deal with more complex operations on data.

---

class: middle inverse

.font50[Solution 2: Creating Objects]

---

class: middle

```{r}
numbers <- c(4, 8, 16)
numbers
```

--

<hr>

```{r}
avg_number <- mean(numbers)
avg_number
```

--

<hr>

```{r}
round(avg_number)
```

---

class: middle

**Problem with creating many objects**

We will end up with too many objects in `Environment`. 

---

class: middle inverse

.font50[Solution 3: The (forward) Pipe Operator %>% ]

---

class: middle

.font75[Shortcut: <br>Ctrl (Command) + Shift + M]

---
class: middle

.pull-left[
```{r}
c(4, 8, 16) %>% 
  mean() %>% 
  round()
```
]

.pull-right[

Combine 4, 8, and 16 `and then`  
Take the mean   `and then`  
Round the output

]


---

class: middle

The output of the first function is the first argument of the second function.

---

Do you recall composite functions such as $f \circ g(x)$?

--

Now we have $f \circ g \circ h (x)$ or `round(mean(c(4, 8, 16)))`

--

.pull-left[

```{r eval = FALSE}
h(x) %>% 
  g() %>% 
  f()
```

]

.pull-right[

```{r eval = FALSE}
c(4, 8, 16) %>% 
  mean() %>% 
  round()
```

]
