---
title: "Scatterplots"
author: "Zahra Moslemi"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css", "https://cdnjs.cloudflare.com/ajax/libs/font-awesome/5.15.3/css/all.min.css"]
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "cols_macro.js"
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"

---

class: title-slide

```{r echo = FALSE, warning=FALSE}
library(fabricerin)

```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

]


---
### Scatterplot

- A Scatterplot displays the relationship between 2 numeric variables. 

- Each dot represents an observation. Their position on the X (horizontal) and Y (vertical) axis represents the values of the 2 variables. 

- Using ggplot2, scatterplots are built thanks to the **geom_point**


---
 
- The _plot()_ function of R allows to build a scatterplot. Both numeric variables of the input dataframe must be specified in the _x_ and _y_ argument.

-  Let’s consider the correlation between the right hippocampus volume of the subjects versus their left hippocampus volume.

```{r warning=FALSE, message=FALSE, include=FALSE}
library(tidyverse)

data <- read_csv("https://raw.githubusercontent.com/COSMOS-DataScience/slides/main/data/alzheimer_data.csv")

attach(data)
```

```{r, fig.height=4.9, fig.align='center'}
plot(lhippo,rhippo)
```

---

### Customizations
- cex: circle size
- xlim and ylim: limits of the X and Y axis
- pch: [shape of markers.](http://www.sthda.com/english/wiki/r-plot-pch-symbols-the-different-point-shapes-available-in-r)
- xlab and ylab: X and Y axis labels
- col: [marker color](https://bit.ly/colors-r)
- main: chart title

---

```{r, fig.height=4.4, fig.align='center'}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col="cadetblue3",
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
```

---


### Map the marker color to a categorical variable

```{r, fig.height=4.4, fig.align='center', fig.show='hide'}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=2, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
```
---
```{r, fig.height=6, fig.align='center', echo=FALSE}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
```
---

### Add legend to a plot 
- Legend function in R adds legend box to the plot. 
- legend() function in R makes graph easier to read and interpret in better way.

```{r, fig.height=4.4, fig.align='center', fig.show='hide'}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
legend("topleft"
       ,legend = c("male","female"),
       pch = 19, col = factor(levels(factor(female)))
       )
```

---
```{r, fig.height=4.4, fig.align='center',  echo=FALSE}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
legend("topleft"
       ,legend = c("male","female"),
       pch = 19, col = factor(levels(factor(female)))
       )
```
---

## Add fit lines

```{r, fig.height=4.4, fig.align='center', fig.show='hide'}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col="black",
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
abline(lm(rhippo~lhippo), col="red")
```

---

```{r, fig.height=4.4, fig.align='center', echo = FALSE}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col="black",
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
abline(lm(rhippo~lhippo), col="red")
```
---

```{r, fig.height=4.4, fig.align='center', fig.show='hide'}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
abline(lm(rhippo[female=="1"]~lhippo[female=="1"]), col="red")
abline(lm(rhippo[female=="0"]~lhippo[female=="0"]), col="blue")
```
---
```{r, fig.height=6, fig.align='center', echo=FALSE}
plot(lhippo,rhippo,
     xlim=c(0,6) , ylim=c(0,6), 
     pch=19, 
     cex=1, 
     col=factor(female),
     xlab="left hippocampus volume", ylab="Right hippocampus volume",
     main="A simple scatterplot"
     )
abline(lm(rhippo[female=="1"]~lhippo[female=="1"]), col="red")
abline(lm(rhippo[female=="0"]~lhippo[female=="0"]), col="blue")

```



---
- The scatterplot( ) function in the car package offers many enhanced features, including fit lines, marginal box plots and conditioning on a factor. Each of these features is optional.

```{r error = FALSE, message=FALSE,warning=FALSE, include=FALSE}
library(car)
```

---

```{r fig.height=3.9, fig.align='center', warning=FALSE}
library(car)
scatterplot(rhippo ~ lhippo | female, 
            boxplots = "xy",
            xlab="left hippocampus volume", ylab="Right hippocampus volume",
            main="Enhanced Scatter Plot")
```

---
### Basic Scatterplot Matrix

```{r fig.show='hide'}
ran <- sample(1:2700,30)
new_data <- data[ran,]

pairs(~naccicv + csfvol + lhippo + rhippo + frcort,
      data=new_data,
      main="Simple Scatterplot Matrix")
```
---

```{r fig.align='center', echo=FALSE}
ran <- sample(1:2700,30)
new_data <- data[ran,]

pairs(~naccicv + csfvol + lhippo + rhippo + frcort, data=new_data,
   main="Simple Scatterplot Matrix")
```
