---
title: "Exploring Alzheimer's Data with R"
author: "Brian Schetzsle"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css", "https://cdnjs.cloudflare.com/ajax/libs/font-awesome/5.15.3/css/all.min.css"]
    lib_dir: libs
    seal: false
    nature:
      beforeInit: "cols_macro.js"
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"

---

class: title-slide

```{r include=FALSE}
library(fabricerin)
library(magrittr)
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`

Adapted from slides by Sam Behseta

]

---

class: middle

Read in the Alzheimer's data and load `tidyverse` package

```{r eval = FALSE}
library(tidyverse)
AD <- readr::read_csv("./data/alzheimer_data.csv")
```

```{r include = FALSE}
library(tidyverse)
AD <- read.csv("./data/alzheimer_data.csv")
```

---

class: middle

Inspect the data

```{r}
dim(AD)
```

---

class: middle

Inspect the data

```{r}
glimpse(AD)
```

---

class: middle

### Subsetting the data

```{r}
AD_subset <- AD %>% select(id,diagnosis,educ)
head(AD_subset)
```

---

class: middle

### Recoding a variable

```{r}
AD_subset <- AD_subset %>% 
  mutate(diagnosis = as.factor(diagnosis))
summary(AD_subset$diagnosis)
```

---

class: middle

Visualize the distribution of a categorical variable

```{r fig.height = 3, fig.align="center"}
AD_subset %>% 
  ggplot() +
  geom_bar(aes(x = diagnosis)) +
  scale_x_discrete(breaks = c(0, 1, 2), 
                   labels = c("Normal", "Low", "Severe"))  +
  labs(x = "Diagnosis", y = "Frequency", 
       title = "Frequency of Diagnoses in the Alzheimer's Data")
```

---

class: middle

The `naccicv` feature is a numerical variable representing the overall volume of the brain. For the patients with more progressive AlZheimer's the volume is typically smaller.

Visualize the distribution of this numerical (continuous) variable.

```{r fig.height = 2, fig.align="center", message=FALSE}
AD %>% 
  select(naccicv) %>% 
  ggplot() +
  geom_histogram(aes(x = naccicv))
```

---

class: middle

A boxplot can also be an effective way of quickly visualizing the distribution of a continuous variable.

```{r fig.height = 2, fig.align="center"}
AD %>% 
  ggplot(aes(x = "", y = naccicv)) +
  geom_boxplot() +
  labs(x = "NACCICV", y = "Units", 
       title = "Distribution of Intracranial Volume")
```

---

class: middle

Are you curious about what the quartile values are displayed in the boxplot? You can retrieve those with the `quantile()` function in R.

```{r}
quantile(AD$naccicv)
```

---

### Practice

- Select a continuous variable from the Alzheimer's data and generate a histogram and a boxplot
- Select a categorial variable and a continuous variable; generate boxplots for each level of the categorical variable
- Select two continuous variables; generate a scatterplot using the `geom_point()` function